---
title: "Untitled"
author: "Brook Andreas"
date: "July 18, 2018"
output: html_document
---
## Github access 

## https://github.com/brook0616/MSDS6306-Doing-Data-Science-HomeWork-/blob/master/BrookAndreasHomework%205%20.Rmd 

### Questions

### Backstory: Your client is expecting a baby soon.  However, he is not sure what to name the child.  Being out of the loop, he hires you to help him figure out popular names.  He provides for you raw data in order to help you make a decision.

### 1.	Data Munging : Utilize yob2016.txt for this question. This file is a series of popular children's names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.

#### a.	First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you're dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.
```{r}
install.packages("data.table",repos = "http://cran.us.r-project.org")
```

```{r}
## importing data from the local diretory , you can also import directly from the dropbox site using RCurl library 
URL <- "https://www.ssa.gov/oact/babynames/state/namesbystate.zip"
temp <- tempfile()
download.file(URL,temp)

### I have downloaded the data to my current working directory 
 my_data <- read.table("C:\\Users\\brook\\Desktop\\SMU courses\\#MSDS 6306 Doing data science\\Homework\\Assignment 05\\yob2016.txt",header = FALSE, sep = " ",fill=FALSE, strip.white=TRUE)
head(my_data)
## Please install the package and import the library 
library(tidyr)
library(dplyr)
## Split the coloumn names 
yob2016 <- my_data %>% separate(V1, c("FirstName", "Gender", "NumberOfChildren16"))
yob2016
head(yob2016)
str(yob2016)
## Change the NumberOfChildren Coloumn to numeric  from char and check 
yob2016$NumberOfChildren16 <- as.numeric(as.character(yob2016$NumberOfChildren))
str(yob2016)
```

#### b.	Display the summary and structure of df
```{r}
str(yob2016)
summary(yob2016)
```
#### c.	Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name.  Write an R command to figure out which name it is and display it.


```{r}
# call the SQL library after installing the package 
library(sqldf)
NameReport <- sqldf("SELECT FirstName  FROM yob2016 WHERE FirstName like '%yyy' ")
head(NameReport)
## found only one data to meet client claim ,Not thousands!!
```
#### d.	Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant.  Save the remaining dataset as an object: y2016 

```{r}
## removing duplicate record if any 

library(data.table);
y2016 <-unique(yob2016, by = "FirstName")
dim(y2016)
yob2016 <- na.omit(yob2016)
```

### 2. Data Merging : Utilize yob2015.txt for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

#### a.	Like 1a, please import the .txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to y2015.

```{r}
## importing data from the local diretory , you can also import directly fro the dropbox site using RCurl library 
y2015 <- read.table("C:\\Users\\brook\\Desktop\\SMU courses\\#MSDS 6306 Doing data science\\Homework\\Assignment 05\\yob2015.txt",header = FALSE, sep = " ",fill=FALSE, strip.white=TRUE);
head(y2015);
library(tidyr);
library(dplyr);
y2015 <- my_data %>% separate(V1, c("FirstName", "Gender","NumberOfChildren15"))
head(y2015)
dim(y2015)
```

```{r} 
## Change the NumberOfChildren15 Coloumn to numeric 
y2015$NumberOfChildren15 <- as.numeric(as.character(y2015$NumberOfChildren15))
str(y2015)

```

  
#### b.	Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.
```{r}
tail(y2015,10)
```
#### c.	Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.
```{r}
dim(y2015)
dim(y2016)
## merges these identical csv files by two common coloumns 
Merged_final <-merge(y2016, y2015, by=c("FirstName","Gender"))
str(Merged_final)
head(Merged_final)


## Checks if there is NA in your data set
NACheck <- y2016[complete.cases(y2016), ]
NACheck <- y2015[complete.cases(y2015), ]
## Code checks if there is NA and lists all rows with NON NA Value.As its dimension is the same as the original data frame , t seems there is no NA 
sqldf::sqldf("select FirstName from Merged_final where NumberOfChildren15='N.A'or 'NA'")
sqldf::sqldf("select FirstName from Merged_final where Gender='N.A'or 'NA'")
sqldf::sqldf("select FirstName from Merged_final where FirstName='N.A'or 'NA'")

## Checks if ther is any NA in the Merged final file 

dim(NACheck)

```
### 3.	Data Summary : Utilize your data frame object final for this part.
#### a.	Create a new column called "Total" in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?


```{r}

## creates a new computed coloumn called Total 
Merged_final<- within(Merged_final, Total <- NumberOfChildren15 +
NumberOfChildren16)


## checks MAx value from the new merged data set 
max(Merged_final$Total,na.rm = TRUE)
## 38,828 ppl were given the most popular name 

## Removes the two numeric coloumns from the final merge 
Merged_final <- subset(Merged_final, select = -c(NumberOfChildren15,NumberOfChildren16))
```

#### b.	Sort the data by Total.  What are the top 10 most popular names?

```{r}
library(plyr);
## Sorts data in descending order 
Arranged_Final<-arrange(Merged_final,desc(Total),Gender)
head(Arranged_Final,10)
Arranged_Final %>% top_n(10)
```
#### c.	The client is expecting a girl!  Omit boys and give the top 10 most popular girl's names.

```{r}
library(sqldf)
Female_Top <- sqldf("select  * from Arranged_Final where Gender='F'")
Female_Top10 <- head(Female_Top,10)
```
#### d.	Write these top 10 girl names and their Totals to a CSV file.  Leave out the other columns entirely.omit the Gender Coloumn 
```{r}
## omit the Gender Coloumn , 
Female_Top10 <- subset(Female_Top, select = -c(Gender))
write.csv(Female_Top10, file = "Female_Top10.csv",row.names=TRUE, na="")
getwd()
list.files()

```

### 4.	Upload to GitHub (10 points): Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week).  The Codebook should contain a short definition of each object you create, and if creating multiple files, which file it is contained in.  You are welcome and encouraged to add other files-just make sure you have a description and directions that are helpful for the grader.

```{r}

## how to generate a code book for the entire project by using all the data frames created through out 

library(dataMaid)

## list of data frames created through out the project 

##makeCodebook(y2016)
##makeCodebook(y2015)
##makeCodebook(Merged_final)
##makeCodebook(Arranged_Final)
## makeCodebook(Female_Top)
## All the codebooks Generated are stored in the Git
```


  
